\sysfull

### 2018-04-14 paper/background.tex
editorial changes to background: table 1 and intro passage to sec 2.3.1

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

\subsubsection{Costs of Task-based Models}

%Task-based models incur costs that motivate \sys.

@@

%\begin{figure}
%\centering
	%\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
	%\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
	%\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
	%\label{fig:framEnergy}
%\end{figure}

To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations. We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching. Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. In an intermittent device, energy cost corresponds to decreased run time between power failures. The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.

>>>>

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}	& Registers + Stack     \\
		DINO~\cite{dino}	& Registers + WAR variables \\
		Chain~\cite{chain}	& PC + channel data\\
		Alpaca~\cite{alpaca}	& PC + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

\subsubsection{Costs of Task-based Models}
\label{sec:cost_task-based}

Despite its advantages, the task-based programming model for intermittently-powered computing devices has its limitations. Two major ones being: {\em memory overhead} and {\em inflexibility to changing energy conditions}.

@@

To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations. We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching. Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. This observation is also consistent with the TI specification~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek} In an intermittent device, energy cost corresponds to decreased run time between power failures. The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.
